Hydrodynamics of two vesicles interacting via a Van der Waals potential

### 2018-12-25 vesicle_code/docs/adhesion/adhesion.tex
fixed bug in adhesion energy term

We investigate rheology of suspensions by computing the effective
viscosity of a doublet and compare it to the effective viscosity of a
single tank-treading vesicle. The effective viscosity is defined as the
viscosity of a homogeneous Newtonian fluid with the same energy
dissipation per macroscopic element of fluid.  In a simple shear flow,
the ambient velocity field is given by $\uu(\xx) = \chi(y,0)$, where
$\chi$ is the shear rate, and the intrinsic viscosity is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 T} \int_{T_i}^{T_e} 
  \langle \sigma_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \sigma \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds.
\end{align*}
Here, $[\mu]$ is the intrinsic viscosity, $|phi$ is the area fraction of
vesicles, $\sigma$ is the stress due to the vesicles, and $\langle \cdot
\rangle$ is the spatial average.

\begin{figure}[htp]
  \includegraphics[scale=1]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} The intrinsic viscosity
  of a tank treading vesicle (blue) and a doublet (red).  The shear rate
is $\chi = 0.5$, the Hamaker constant is $\sigma = 0.7$, and the
adhesion length scale is $\delta = 0.4$.  The length of the vesicles at
each of the reduced areas is fixed, while the area changes with
differing reduced areas.}
\end{figure}

@@

We now decompose the perturbation into normal and tangential components
as $\delta \xx_1 = \epsilon \yy(s) = \epsilon(u\tt + v\nn)$. The
perturbed arclength term, to leading order, is
\begin{align*}
  \|\tilde{\xx}'_1\| \approx 1 + \epsilon(u_s + \nu\kappa),
\end{align*}
where $\kappa$ is the curvature.  To leading order, inextensible
perturbations satisfy $u_s + \kappa v = 0$, so the arclength term of
$\gamma_1$ and $\tilde{\gamma}_1$ are identical to leading order.
Therefore,
\begin{align*}
  \delta U_1 &= \int_{\gamma_1} \int_{\gamma_2} \left(
  \phi(\|\xx_1 + \epsilon \yy  - \xx_2\|) - \phi(\|\xx_1 - \xx_2\|)
  \right) ds_{\xx_2} ds_{\xx_1} \\
  &\approx \epsilon \int_{\gamma_1} \int_{\gamma_2}
  \nabla \phi (\|\xx_1 - \xx_2\|) \cdot \yy ds_{\xx_2} ds_{\xx_1}
\end{align*}
and the adhesive force applied by vesicle 2 on vesicle 1 is
\begin{align*}
  \int_{\gamma_2}\nabla \phi(\|\xx_1 - \xx_2\|)ds_{\xx_2} = 
  -4 \sigma \delta^2 \int_{\gamma_2}
  \frac{\xx_1 - \xx_2}{\|\xx_1 - \xx_2\|^2} 
  \left(\delta^2 - \|\xx_1 - \xx_2\|^2 \right) ds_{\xx_2}.
\end{align*}
A similar expression holds for the adhesive force applied by vesicle 1
on vesicle 2, and equation~\eqref{eqn:adhesionForce} gives the adhesive
force for a suspension of $M$ vesicles.

>>>>

We investigate rheology of suspensions by computing the effective
viscosity of a doublet and compare it to the effective viscosity of a
single tank-treading vesicle. The effective viscosity is defined as the
viscosity of a homogeneous Newtonian fluid with the same energy
dissipation per macroscopic element of fluid.  In a simple shear flow,
the ambient velocity field is given by $\uu(\xx) = \chi(y,0)$, where
$\chi$ is the shear rate, and the intrinsic viscosity is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 T} \int_{T_i}^{T_e} 
  \langle \sigma_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \sigma \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds.
\end{align*}
Here, $[\mu]$ is the intrinsic viscosity, $\phi$ is the area fraction of
vesicles, $\sigma$ is the stress due to the vesicles, and $\langle \cdot
\rangle$ is the spatial average.

\begin{figure}[htp]
  \includegraphics[scale=1]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} The intrinsic viscosity
  of a tank treading vesicle (blue) and a doublet (red).  The shear rate
is $\chi = 0.5$, the Hamaker constant is $\sigma = 0.7$, and the
adhesion length scale is $\delta = 0.4$.  The length of the vesicles at
each of the reduced areas is fixed, while the area changes with
differing reduced areas.}
\end{figure}
\todo[inline]{Do the smaller reduced area doublets since the intrinsic
viscosity seemed to increase as the reduced area was decresed further}
\todo[inline]{Decompose the red curve as the intrinsic viscosity due to
adhesion and due to the bending/tension}

@@

We now decompose the perturbation into normal and tangential components
as $\delta \xx_1 = \epsilon \yy(s) = \epsilon(u\tt + v\nn)$. The
perturbed arclength term, to leading order, is
\begin{align*}
  \|\tilde{\xx}'_1\| \approx 1 + \epsilon(u_s + \nu\kappa),
\end{align*}
where $\kappa$ is the curvature.  To leading order, inextensible
perturbations satisfy $u_s + \kappa v = 0$, so the arclength term of
$\gamma_1$ and $\tilde{\gamma}_1$ are identical to leading order.
Therefore,
\begin{align*}
  \delta U_1 &= \int_{\gamma_1} \int_{\gamma_2} \left(
  \phi(\|\xx_1 + \epsilon \yy  - \xx_2\|) - \phi(\|\xx_1 - \xx_2\|)
  \right) ds_{\xx_2} ds_{\xx_1} \\
  &\approx \epsilon \int_{\gamma_1} \int_{\gamma_2}
  \nabla \phi (\|\xx_1 - \xx_2\|) \cdot \yy ds_{\xx_2} ds_{\xx_1}
\end{align*}
and the adhesive force applied by vesicle 2 on vesicle 1 is
\begin{align*}
  \int_{\gamma_2}\nabla \phi(\|\xx_1 - \xx_2\|)ds_{\xx_2} = 
  -4 \sigma \delta^2 \int_{\gamma_2}
  \frac{\xx_1 - \xx_2}{\|\xx_1 - \xx_2\|^6} 
  \left(\delta^2 - \|\xx_1 - \xx_2\|^2 \right) ds_{\xx_2}.
\end{align*}
A similar expression holds for the adhesive force applied by vesicle 1
on vesicle 2, and equation~\eqref{eqn:adhesionForce} gives the adhesive
force for a suspension of $M$ vesicles.
